The Place of the Local Group in the Cosmic Web

### 2014-08-05 paper/main.tex
small modifications in the results section

\newcommand{\lcdm}{$\Lambda$CDM }
\newcommand{\mpc}{\rm{Mpc}}
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\begin{abstract}
In this Letter we explore the characteristics of the Local Group (LG)
location in the cosmic web using a cosmological simulation in the
$\Lambda$CDM cosmology.
We use the Hessian of the gravitational potential to define the cosmic web, 
where regions are classified as either peak, sheet, filament or
void. 
The LG in the simulations are split into two samples.
The first is a general sample composed by halo pairs with
similar masses and isolation criteria as observed in the LG. 
The second is a subset with additional kinematic constaints.
We find that the pairs in the LG sample with all constraints are:
i) Preferentially located in filaments and sheets,
ii) located in in a narrow range of local overdensity $0<\delta<4$, 
web ellipticity $0.1<e<1.0$ and prolateness $-0.5<p<0.5$. 
iii) Clearly aligned with the cosmic web, in particular for pairs in
filaments/sheets the orbital angular momentum tends to be
perpendicular to the filament direction or the sheet plane. A stronger
alignment is present for the vector linking the two halos, which lies
along the filament or the sheet plane.  
We show that the first and second results are expected trends with the
LG total mass. 
However, the strong alignment with the cosmic web cannot be explained
in the same way.

@@

%this paper.
In this Letter we study the large scale environment of LG 
analogues in the context of $\Lambda$CDM. We use the Bolshoi
simulation to explore in what structures they reside and if there is
any correlation or alignment with the cosmic web. The large scale
Environment is defined by the cosmic web components identified by
\citet{Tweb}, and we use the LG analogues computed by
\citet{lganalogues}. We pay spetial attention to quantify the kind of
environment that hosts LG pairs and their alignments with respect to
the prefered directions defined by the T-web.

This Letter is organized as follows. In Section \ref{sec:simulation}
we present the N-body cosmological simulation and the algorithm to
define the cosmic web, next in \ref{sec:lg_analogues} we describe the
sample of LG analogues extracted from the simulation. In
Section \ref{sec:results} we presents our results to continue with a
discussion about their implications for our Local Group in Section
\ref{sec:discussion} to finally conclude in Section
\ref{sec:conclusions}.

@@

\cite{Tweb} performed a detailed study for the topology of the
cosmic web and its visual counterpart as a function of the parameter
$\lambda_{\rm th}$. 
They found that reasonable results in terms of the
volume fraction occupied by voids, the visual inspection and the halo
populations in each web type can be reached by values of $0.2<\lambda_{\rm
th}<0.4$. 
In this Letter we choose the value of $\lambda_{\rm
  th}=0.3$ to proceed with our analysis. 
This is only relevant to the classification of the simulation into web
elements. Other results are completeley independent of this
choice. Nevertheless we have checked that the main conclusions of this
work do not depend on the choise of $\lambda_{\rm th}$.

The algorithm to compute the potential is grid based. First we
interpolate the density into a cubic grid with a Cloud-In-Cell (CIC)
scheme and smooth it with a gaussian kernel. 
Then we obtain the
gravitational potential using FFT methods and use finite differences
to compute the Hessian at every point in the grid. 
In our case we have
used a grid size on and a gaussian smoothing with two times larger as
the typical separation between the two halos in the Local Group. 
The
purpose of this choice is to have both halos in the pair a common
environment. In this paper we use a grid spacing of $s=0.97$ \hMpc,
corresponding to a $256^3$ grid in the Bolshoi volume. 
The scale for
the gaussian smoothing uses the same value.

@@

We are also interested in measuring the alignment of the LG halos with
respect to the cosmic web. To this end we characterize each LG pair by
the vector that is normal to the orbital plane, which is defined by
the position and velocities of each halo with respect to the center of
mass.

@@

The LG analogues or General sample in this paper are pairs selected in relative isolation,
and in a wide range of masses from  $M_{200c}=5 \times 10^{10}$ \msun $ $ to $ 5 \times 10^{13}$ \msun. 
Isolation criteria includes a pair closer than $1.3$\mpc, and with no massive neighbors within $5$\mpc.
In addition we require that pairs have no Virgo-like neighbor halo with mass $M_{200c}>1.5 \times 10^{14}\ \rm M_{\odot}$ within $12$ Mpc. 
We have $5480$ pairs under these general criteria.
A full description of the selection criteria can be found in \citet{lganalogues,sat}.

We also define two subsamples more closely related to the MW-M31 
dynamics  according to the tolerance in additional constraints:
A sample named $2\sigma$, correspondig to LG analogues constrained by
two times the observational errors in the orbital values (radial
velocity, tangential velocity, and separation), and a more relaxed
sample named $3\sigma$ for LG analogues constrained by three times
observational errors accordingly. The number of pairs in each sample
is $46$ and $120$ respectively, notice we have less pairs than in
\citet{lganalogues} results, since we removed pairs which are too
close at $z=0$, i.e. their virial radii overlaps, also we removed a
couple pairs that merged or change their mass more than $20\%$ at
present time since they were detected at $z<0.1$.

%We also have a third general sample that we use as a control for our
%results.

@@

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}\hline\hline
Sample & Peak & Filament & Sheet & Void\\
       & $n$ (\%) & $n$ (\%) & $n$ (\%) & $n$ (\%) \\\hline
2$\sigma$ & 4 (8.7) & 24 (52.2) &  17 (36.7) & 1 (2.2)\\
3$\sigma$ & 10 (8.3) & 58 (48.3) & 47 (39.2) & 5 (4.2)\\  
General ($12.1<\log_{10} M_{\rm h}/\Msun<12.3$)& 8 (1.4) & 334 (55.5) & 259
(43.0) & 1 (0.1)\\
General & 1312 (23.9) & 1472 (26.9) & 1769 (32.3) & 927 (16.9)\\

\hline\hline
\end{tabular}
\caption{
Number of pairs in the four different kinds of environments. In
parethesis the same number as a percentage of the
total population.
\label{table:web_type}}
\end{center}
\end{table*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{histogram_mass_distro.pdf}
\caption{Mass distribution of pairs in the different environments
for the General sample.
Dashed regions correspond to the mass ranges of $2\sigma$ and $3\sigma$ samples.
\label{fig:median_fraction}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{alignments_e1_n_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e2_n_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e3_n_all_environments.pdf} 
\end{center}
\caption{Cumulative distributions for the alignment between the normal vector to the pair orbit
$\hat{n}$, and the eigenvectors defining the cosmic web.
    \label{fig:alignment}}  
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{alignments_e1_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e2_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e3_r_all_environments.pdf} 
\end{center}
\caption{Cumulative distributions for the alignment between the unit
  vector linking the two halos, $\hat{r}$, and the eigenvectors
  defining the cosmic web.
    \label{fig:alignment_r}}  
\end{figure}

@@

The situation in the restricted $2\sigma$ and $3\sigma$ samples is very
different. By large the LGs in these samples are located in
filaments and sheets. In both samples, $\sim 50\%$ of the
pairs can be found in filaments while $\sim 40\%$ are in
in sheets. These absolute numbers in each environment for each sample
are presented in Table \ref{table:web_type}.

This difference between the general and the restricted samples is mostly due 
the mass ranges covered by each sample. 
In \citet{lganalogues} the mass range covered by $2\sigma$ and $3\sigma$ 
is very narrow and it is used to constraint the LG mass, then we show 
also in table \ref{table:web_type} a subsample of the General sample
having a similar mass range to $2\sigma$ and $3\sigma$, and it reproduces 
similar environment fractions.

A more clear view of this environment dependency and pair mass is shown in
figure \ref{fig:median_fraction}, where we show the mass distribution of pairs
in the different environments, then high mass pairs tend to be located in 
peaks and filaments, and less massive ones in voids and sheets.
Dashed regions represent the $68\%$ confidence intervals of the mass 
distributions of $2\sigma$ and $3\sigma$ samples.

@@

\subsection{Overdensity, Ellipticity and Anisotropy}

We now describe the preferred place of the LG samples in terms of the
overdensity, ellipticiy and anisotropy as defined in Section
\ref{sec:simulation}.

Figure \ref{fig:median_overdensity} shows dependency of overdensity,
ellipciticy, and prolateness on pair mass for the different samples.
General sample is represented by the solid lines with the associated 
errors covered by the shaded region.

Left panel shows the overdensity dependence on mass where higher mass
pairs are located in high density regions.
The $2\sigma$ and $3\sigma$ samples having a narrower mass range as 
shown in previous figure, are consequently located within a narrower 
range of overdensities $0.0<\delta<4.0$ peaking at $\delta \sim 1$. 
This is also consistent with the fact that these samples are mostly 
found in filaments and sheets. 
Average overdensity of $2\sigma$ and $3\sigma$ samples show no 
significant difference from the General sample within the same mass 
range.

%Figure \ref{fig:density} shows that the range of overdensities for the
%general pair sample spans almost two orders of magnitude from
%underdense regions $\delta<0$ to high density regions $\delta\sim 30$,
%with a distribution peaking around $\delta \sim 0$. In contrast the
%$2\sigma$ and $3\sigma$ samples are located within a narrower range of
%overdensities $0.0<\delta<4.0$ peaking at $\delta \sim 1.5$. This is
%consistent with the fact that these samples are mostly found in
%filaments and sheets.

Middle and right panels show web ellipciticy and absolute prolateness
dependence on mass. Again we noticed that within the same mass range,
the $2\sigma$ and $3\sigma$ average ellipciticy and prolateness does not
differ significantly from General sample.
For the $2\sigma$ and $3\sigma$ samples most of the pairs are located
in a narrow range for ellipticities in the range $0.1<e<1.0$, and
prolateness $0.5<p<0.5$.

>>>>

\newcommand{\manuscript}{paper }
\newcommand{\lcdm}{$\Lambda$CDM }
\newcommand{\mpc}{\rm{Mpc}}
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\begin{abstract}
We explore the characteristics of the Local Group (LG)
location in the cosmic web using a cosmological simulation in the
$\Lambda$CDM cosmology.  We use the Hessian of the gravitational
potential to define the cosmic web, where regions are classified as
either peak, sheet, filament or void.  The LG in the simulations are
split into two samples.  The first is a general sample composed by
halo pairs with similar masses and isolation criteria as observed in
the LG.  The second is a subset with additional kinematic constaints.
We find that the pairs in the LG sample with all constraints are: i)
Preferentially located in filaments and sheets, ii) located in in a
narrow range of local overdensity $0<\delta<4$, web ellipticity
$0.1<e<1.0$ and prolateness $-0.5<p<0.5$.  iii) Clearly aligned with
the cosmic web, in particular for pairs in filaments/sheets the
orbital angular momentum tends to be perpendicular to the filament
direction or the sheet plane. A stronger alignment is present for the
vector linking the two halos, which lies along the filament or the
sheet plane.  We show that the first and second results are expected
trends with the LG total mass.  However, the strong alignment with the
cosmic web cannot be explained in the same way.

@@

%this paper.
In this \manuscript we study the large scale environment of LG 
analogues in the context of $\Lambda$CDM. We use the Bolshoi
simulation to explore in what structures they reside and if there is
any correlation or alignment with the cosmic web. The large scale
Environment is defined by the cosmic web components identified by
\citet{Tweb}, and we use the LG analogues computed by
\citet{lganalogues}. We pay spetial attention to quantify the kind of
environment that hosts LG pairs and their alignments with respect to
the prefered directions defined by the T-web.

This \manuscript is organized as follows. In Section \ref{sec:simulation}
we present the N-body cosmological simulation and the algorithm to
define the cosmic web, next in \ref{sec:lg_analogues} we describe the
sample of LG analogues extracted from the simulation. In
Section \ref{sec:results} we presents our results to continue with a
discussion about their implications for our Local Group in Section
\ref{sec:discussion} to finally conclude in Section
\ref{sec:conclusions}.

@@

\cite{Tweb} performed a detailed study for the topology of the
cosmic web and its visual counterpart as a function of the parameter
$\lambda_{\rm th}$. 
They found that reasonable results in terms of the
volume fraction occupied by voids, the visual inspection and the halo
populations in each web type can be reached by values of $0.2<\lambda_{\rm
th}<0.4$. 
In this \manuscript we choose the value of $\lambda_{\rm
  th}=0.3$ to proceed with our analysis. 
This is only relevant to the classification of the simulation into web
elements. Other results are completeley independent of this
choice. Nevertheless we have checked that the main conclusions of this
work do not depend on the choise of $\lambda_{\rm th}$.

The algorithm to compute the potential is grid based. 
First we interpolate the density into a cubic grid with a
Cloud-In-Cell (CIC) scheme and smooth it with a gaussian kernel. 
Then we obtain the gravitational potential using FFT methods and use finite differences
to compute the Hessian at every point in the grid. 
In our case we have used a grid size on and a gaussian smoothing with
two times larger as the typical separation between the two halos in the Local Group. 
The purpose of this choice is to have both halos in the pair a common
environment. 
In this \manuscript we use a grid spacing of $s=0.97$ \hMpc,
corresponding to a $256^3$ grid in the Bolshoi volume. 
The scale for the gaussian smoothing uses the same value.

@@

We also measure the alignment of the LG halos with respect to the
cosmic web defined by the eigenvector.
To this end we characterize each LG pair by two vectors. 
The first is $\hat{n}$, the vector marking the axis along the orbital angular
momentum of the pair, normal to its orbital plane; the second is
$\hat{r}$, the vector that connects the halos in the pair. 
We quantify the alignment using the absolute value of the cosinus of
the angle between the two vectors of interest $\mu=|\hat{e}_i \cdot
\hat{n}|$ or $\mu=|\hat{e}_i\cdot \hat{r}|$, where $i=1,2,3$.

@@

The LG analogues or General Sample (GS) in this paper are pairs selected in
relative isolation, and in a wide range of masses from  $M_{200c}=5
\times 10^{11}$ \msun $ $ to $ 5 \times 10^{13}$ \msun.  
Isolation criteria includes a pair closer than $1.3$\mpc, and with no
massive neighbors within $5$\mpc.
In addition we require that pairs have no Virgo-like neighbor halo
with mass $M_{200c}>1.5 \times 10^{14}\ \rm M_{\odot}$ within $12$
Mpc.  
We have $5480$ pairs under these general criteria. 
A full description of the selection criteria can be found in
\citet{lganalogues,sat}.

We also define two subsamples more closely related to the MW-M31 
dynamics according to the tolerance in additional constraints.
A sample named $2\sigma$, correspondig to LG analogues constrained by
two times the observational errors in the orbital values (radial
velocity, tangential velocity, and separation), and a more relaxed
sample named $3\sigma$ for LG analogues constrained by three times
observational errors accordingly. 
The number of pairs in each sample is $46$ and $120$ respectively,
notice we have less pairs than in \citet{lganalogues} results, since
we removed pairs which are too close at $z=0$, i.e. their virial radii
overlaps, also we removed a couple pairs that merged or change their
mass more than $20\%$ at present time since they were detected at
$z<0.1$.

@@

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}\hline\hline
Sample & Peak & Filament & Sheet & Void\\
       & $n$ (\%) & $n$ (\%) & $n$ (\%) & $n$ (\%) \\\hline
2$\sigma$ & 4 (8.7) & 24 (52.2) &  17 (36.7) & 1 (2.2)\\
3$\sigma$ & 10 (8.3) & 58 (48.3) & 47 (39.2) & 5 (4.2)\\  
General & 1312 (23.9) & 1472 (26.9) & 1769 (32.3) & 927 (16.9)\\
General ($12.1<\log_{10} M_{\rm LG}/\Msun<12.3$)& 8 (1.4) & 334 (55.5) & 259
(43.0) & 1 (0.1)\\
\hline\hline
\end{tabular}
\caption{
Number of pairs in the four different kinds of environments for each
of the three samples presented in Section \ref{sec:lg_analogues}. 
In parethesis the same number as a percentage of the total population.
The last line in the table corresponds to the general sample with an
additional mass cut for the total pair mass.  
\label{table:web_type}}
\end{center}
\end{table*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{histogram_mass_distro.pdf}
\caption{Mass distribution of pairs in the different environments
for the general sample.
The shaded regions show the mass ranges of $2\sigma$ and $3\sigma$
samples.  
\label{fig:median_fraction}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{alignments_e1_n_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e2_n_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e3_n_all_environments.pdf} 
\end{center}
\caption{Cumulative distributions for the alignment between the normal
  vector to the pair orbital plane, $\hat{n}$, and the three eigenvectors in
  the Tweb.
    \label{fig:alignment_n}}  
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{alignments_e1_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e2_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e3_r_all_environments.pdf} 
\end{center}
\caption{Cumulative distributions for the alignment between the vector
  linking the two halos in the pair, $\hat{r}$, and the three
  eigenvectors in the Tweb.
    \label{fig:alignment_r}}  
\end{figure}

@@

@@

The situation in the restricted $2\sigma$ and $3\sigma$ samples is very
different. 
By large the LGs in these samples are located in filaments and sheets. 
In both samples, $\sim 50\%$ of the pairs can be found in filaments
while $\sim 40\%$ are in sheets. 
These absolute numbers in each environment for each sample are
presented in Table \ref{table:web_type}.

This difference between the general and the restricted samples is
due the mass ranges covered by each sample. 
In \citet{lganalogues} the mass range covered by $2\sigma$ and $3\sigma$ 
is very narrow and it is used to constraint the LG mass.
We show  in table \ref{table:web_type} that subset of the GS
having a similar mass range to $2\sigma$ and $3\sigma$ reproduces  
similar environment fractions.

Figure \ref{fig:median_fraction} clearly shows the correlation between
environment an total pair mass.
Each line represents the mass distribution of pairs in the four
different environments for the general sample.
High mass pairs tend to be located in  peaks and filaments while less
massive ones in voids and sheets. 
The shaded regions represent the $68\%$ confidence intervals of the mass 
distributions of $2\sigma$ and $3\sigma$ samples.

\subsection{Web Overdensity, Ellipticity and Anisotropy}

We now describe the preferred place of the LG samples in terms of the
web overdensity, ellipticiy and anisotropy as defined in Section
\ref{sec:simulation}.

Figure \ref{fig:median_overdensity} shows dependency of the web overdensity,
ellipciticy, and prolateness on pair mass for the different samples.
GS is represented by the solid lines with the associated 
errors covered by the shaded region. 
The symbols represents the results for the $2\sigma$ and $3\sigma$
samples.  
In all cases it is immediatly clear that the range of values for the
$2\sigma$ and $3\sigma$ samples are completely expected from its mass
dependence.

Left panel shows the overdensity dependence on pair mass. 
Higher mass pairs are located in high density regions.
The $2\sigma$ and $3\sigma$ samples having a narrower mass range as 
shown in previous figure, are consequently located within a narrower 
range of overdensities $0.0<\delta<4.0$ peaking at $\delta \sim 1$. 
This is also consistent with the fact that these samples are mostly 
found in filaments and sheets. 
The average overdensity of $2\sigma$ and $3\sigma$ samples is expected
from the values in the GS within the same mass range.

Middle and right panels show web ellipciticy and absolute prolateness
dependence on mass. Again we noticed that within the same mass range,
the $2\sigma$ and $3\sigma$ average ellipciticy and prolateness does not
differ significantly from GS
For the $2\sigma$ and $3\sigma$ samples most of the pairs are located
in a narrow range for ellipticities in the range $0.1<e<1.0$, and
prolateness $0.5<p<0.5$.

@@
